---
title: "103hw5"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=-1)
```



***
1.   Construct a simple linear regression model for Highway MPG vs Horsepower.
a.   Is the P-value for the model significant? (Quote the P-value in your response.)
yes the P-value is 3.7e-11 < .05 so it is significant. 
b.   What is R2 for the model?
R2 is .3832 
c.   Graph the Plot of Fitted Model and include it in your solutions. Use it to describe the problem with the model.
***
```{r}
library(readxl)
cars <- read_excel("/Users/austinwilson/downloads/93cars.xlsx")
model <- lm(formula=`MPG Highway`~Horsepower,data=cars)
#summary(cars)
#names(cars)
summary(model)

```


```{r}


plot(x=cars$Horsepower,y=cars$`MPG Highway`)
lines(cars$Horsepower, fitted(model), col="blue")
```

***
2.   Construct a quadratic model.
a.   Is the P-value for the model significant? (Quote the P-value in your response.)
P-value is 9.78e-15 so this is statistically significant 
b.   What is the P-value for the second-degree term?
4.8e-6
c.   What is R2 for the model? What is? Which is more appropriate in evaluating the model? 2-adjustedRd.  
.5117, adjusted R-squared is .5009, I will use R-squared because we only have one predictor and adjusted is meaningless. 
Graph the Plot of Fitted Model and include it in your solutions. ***


```{r}
model2 <- lm(formula=`MPG Highway`~Horsepower+I(Horsepower^2),data=cars)
summary(model2)



```




```{r}
plot(x=cars$Horsepower,y=cars$`MPG Highway`)
lines(cars$Horsepower, fitted(model2), col="blue")
```

***
3.   Try constructing higher order models until you are satisfied you have found the best polynomial model. 
a.   Write out your final model or display the analysis window for it so I can see the model you settled on.
The 4th degree polynomial funtion has the nighest adjusted R-squared
b.   What is adjustedR2?
.542
c.   Include the Plot of Fitted Model and Residuals versus X output.
d.   Predict the highway mileage of a car with 150 horsepower. Also construct 95% confidence and prediction intervals for such a car.
***

```{r}
# model3 <- lm(formula=`MPG Highway`~Horsepower+I(Horsepower^2)+I(Horsepower^3),data=cars)
model4 <- lm(formula=`MPG Highway`~Horsepower+I(Horsepower^2)+I(Horsepower^3)+I(Horsepower^4),data=cars)
# model5 <- lm(formula=`MPG Highway`~Horsepower+I(Horsepower^2)+I(Horsepower^3)+I(Horsepower^4)+I(Horsepower^5),data=cars)
# summary(model3)
summary(model4)
# summary(model5)
```


```{r}
plot(x=cars$Horsepower,y=cars$`MPG Highway`)
# lines(cars$Horsepower, fitted(model4), col="blue")
lines(spline(cars$Horsepower, fitted(model4)))
```

```{r}

plot(model4$fitted,model4$resid)

```



```{r}

test = data.frame(
  Horsepower = 150
)
predict(model4,newdata=test,interval='confidence')

```